### 2024-07-23 modern-portfolio-theory.qmd
Built site for gh-pages
20240608 - getting started
20240627 - sessionInfo()

>>>>

## Getting Started {#sec-modernPortfolioTheoryGettingStarted}

### Load Packages {#sec-modernPortfolioTheoryLoadPackages}

```{r}
library("quantmod")
library("fPortfolio")
library("tidyverse")
```

## Overview {#sec-modernPortfolioTheoryOverview}

<https://eng.wealthfront.com/2012/01/17/moneyball-using-modern-portfolio-theory-to-win-your-fantasy-sports-league/> (archived at <https://perma.cc/NE35-G6LR>)

## Download Historical Stock Prices {#sec-downloadStockPrices}

```{r}
symbols <- c(
  "AAPL",  # Apple
  "MSFT",  # Microsoft
  "GOOGL", # Google
  "AMZN",  # Amazon
  "META",  # Meta/Facebook
  "V",     # Visa
  "DIS",   # Disney
  "NKE",   # Nike
  "TSLA")  # Tesla

quantmod::getSymbols(symbols)
```

## Calculate Stock Returns {#sec-calculateStockReturns}

```{r}
prices <- do.call(
  merge,
  lapply(
    symbols,
    function(sym) quantmod::Cl(get(sym))))

returns <- na.omit(
  TTR::ROC(
    prices,
    type = "discrete"))

returns_ts <- timeSeries::as.timeSeries(returns)
```

## Create Portfolio {#sec-createPortfolio}

```{r}
portfolioSpec <- fPortfolio::portfolioSpec()

fPortfolio::setNFrontierPoints(portfolioSpec) <- 1000
```

## Determine the Efficient Frontier {#sec-efficientFrontier}

```{r}
efficientFrontier <- portfolioFrontier(
  returns_ts,
  spec = portfolioSpec)

efficientFrontier
```

```{r}
# Extract the coordinates of individual assets
asset_means <- colMeans(returns)
asset_sd <- apply(returns, 2, sd)

# Add some padding to plot limits (so ticker symbols don't get cut off)
xlim <- range(asset_sd) * c(0.9, 1.1)
ylim <- range(asset_means) * c(0.9, 1.1)

xlim[1] <- 0
ylim[1] <- 0

# Set scientific notation penalty
options(scipen = 999)

plot(
  efficientFrontier,
  which = c(
    1,  # efficient frontier
    3,  # tangency portfolio
    4), # risk/return of individual assets
  control = list(
    xlim = xlim,
    ylim = ylim
  ))

# Add text labels for individual assets
points(
  asset_sd,
  asset_means,
  col = "red",
  pch = 19)

text(
  asset_sd,
  asset_means,
  labels = symbols,
  pos = 4,
  cex = 0.8,
  col = "black")
```

## Identify the Optimal Weights {#sec-portfolioOptimalWeights}

### Tangency Portfolio {#sec-tangencyPortfolio}

The tangency portfolio is the portfolio with the highest Sharpe ratio (i.e., the highest ratio of return to risk).
In other words, it is the portfolio with the greatest risk-adjusted returns.

```{r}
# Find the tangency portfolio (portfolio with the highest Sharpe ratio)
tangencyPortfolio <- fPortfolio::tangencyPortfolio(
  data = returns_ts,
  spec = portfolioSpec)

# Extract optimal weights
tangencyPortfolio_optimalWeights <- fPortfolio::getWeights(tangencyPortfolio)
tangencyPortfolio_optimalWeights

# Output the results
summary(tangencyPortfolio)
```

### Portfolio with Max Return at a Given Risk-Level {#sec-tangencyPortfolio}

```{r}
#| eval: false

# Define target risk levels
targetRisks <- seq(0, 0.3, by = 0.01)

# Initialize storage for optimal portfolios
optimalPortfolios <- list()
optimalWeights_list <- list()

# Find optimal weightings for each target risk level
for (risk in targetRisks) {
  # Create a portfolio optimization specification with the target risk
  portfolioSpec <- fPortfolio::portfolioSpec()
  fPortfolio::setTargetRisk(portfolioSpec) <- risk

  # Solve for the maximum return at this target risk
  optimal_portfolio <- fPortfolio::maxreturnPortfolio(
    returns_ts,
    spec = portfolioSpec)

  # Store the optimal portfolio
  optimalPortfolios[[as.character(risk)]] <- optimal_portfolio

  # Store the optimal portfolio weights with risk level
  optimal_weights <- fPortfolio::getWeights(optimal_portfolio)
  optimalWeights_list[[as.character(risk)]] <- c(RiskLevel = risk, optimal_weights)
}

optimalWeightsByRisk <- dplyr::bind_rows(optimalWeights_list)
optimalWeightsByRisk
```

## Conclusion {#sec-modernPortfolioTheoryConclusion}

::: {.content-visible when-format="html"}

## Session Info {#sec-modernPortfolioTheorySessionInfo}

```{r}
sessionInfo()
```

:::
